---
title: "power_simulation_cross_lagged_panel"
author: "Nico Bast"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

# Power in a cross-lagged panel model (task 3.1)

Performs a power simulation for random intercept cross-lagged panel models
- [reference link](https://www.tandfonline.com/doi/full/10.1080/10705511.2022.2122467)
- [vignette](https://cloud.r-project.org/web/packages/powRICLPM/vignettes/powRICLPM.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("powRICLPM")
require(powRICLPM) #power simulation

require(progressr) #progress bar during simulation
require(furrr) #parallel processing

# see: https://cloud.r-project.org/web/packages/powRICLPM/vignettes/powRICLPM.html

# Check how many cores are available and determine a lower number of cores in parallel processing
parallel_cores<-future::availableCores()-4

```

## 1. Define simulation parameters

- Phi: Standardized autoregressive and cross-lagged effects for the within-unit components of the model. These values are collected in a matrix, with columns representing predictors and rows representing outcomes.
- within_cor: A correlation for the within-unit components.
- ICC: The proportion of variance at the between-unit level (relative to the total variance).
- RI_cor: The correlation between the random intercepts.

If you are unsure if you have specified the Phi matrix as intended, you can use the check_Phi() function to give you a summary of how the effects in your Phi are interpreted.

```{r values}

Phi <- matrix(c(.7, .1, .2, .7), ncol = 2, byrow = T)
# The .2 refers to our standardized cross-lagged effect of interest
check_Phi(Phi)

within_cor <- 0.3
ICC <- 0.5
RI_cor <- 0.3

```
## 2. Perform power simulation

```{r sim, warnings=FALSE}

# Plan powRICLPM analysis to run on 1 core less than number of available cores
plan(multisession, workers = parallel_cores) # For the case of 8 available cores

#power simulation with progress bar
with_progress({ # Subscribe to progress updates

output <- powRICLPM(
  
  target_power = 0.8,
  
  search_lower = 100, #sample size min
  search_upper = 400, #sample size max
  search_step = 50, # step increase
  
  #time_points = c(3, 4, 5),
  time_points = 5, # timepoints used
  
  #taken from parameter definition above
  ICC = ICC,
  RI_cor = RI_cor, 
  Phi = Phi,
  
  within_cor = 0.3, #default medium size
  
  reps = 500 #number of monte carlo simulations
  
)

})

# Revert back to sequential execution of code 
plan(sequential)

```
## 3. inspect output

```{r output}

output

summary(output)

# cross-lagged effect in different simulations
summary(output, parameter = "wB2~wA1") 
plot(output, parameter = "wB2~wA1")

```

# Power in linear mixed model (task 1.1)


