---
title: "power_simulation_cross_lagged_panel"
author: "Nico Bast"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Power in the statistical models of the Emmy-Noether Grant 2024:
- cross-lagged panel model (task 3.1)
- linear mixed model (task 1.1)
- linear model (task 2.1)
- structural equation model (task 2.2)
- bivariate latent growth model with time-invariant covariate (task 3.2)


#power simulation for random intercept cross-lagged panel models

- [reference link](https://www.tandfonline.com/doi/full/10.1080/10705511.2022.2122467)
- [vignette](https://cloud.r-project.org/web/packages/powRICLPM/vignettes/powRICLPM.html)

```{r setup power 1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("powRICLPM")
require(powRICLPM) #power simulation
require(progressr) #progress bar during simulation
require(furrr) #parallel processing

# see: https://cloud.r-project.org/web/packages/powRICLPM/vignettes/powRICLPM.html

# Check how many cores are available and determine a lower number of cores in parallel processing
core_offset<-4 #should be at least 1
parallel_cores<-future::availableCores()-core_offset #cpu core used in parallel processing

```

## 1. Define simulation parameters

- Phi: Standardized autoregressive and cross-lagged effects for the within-unit components of the model. These values are collected in a matrix, with columns representing predictors and rows representing outcomes.
- within_cor: A correlation for the within-unit components.
- ICC: The proportion of variance at the between-unit level (relative to the total variance).
- RI_cor: The correlation between the random intercepts.

If you are unsure if you have specified the Phi matrix as intended, you can use the check_Phi() function to give you a summary of how the effects in your Phi are interpreted.

```{r values}

Phi <- matrix(c(.7, .1, .2, .7), ncol = 2, byrow = T)
# The .2 refers to our standardized cross-lagged effect of interest
check_Phi(Phi)

within_cor <- 0.3
ICC <- 0.5
RI_cor <- 0.3

```
## 2. Perform power simulation

```{r sim, warnings=FALSE}

# Plan powRICLPM analysis to run on 1 core less than number of available cores
plan(multisession, workers = parallel_cores) # For the case of 8 available cores

#power simulation with progress bar
with_progress({ # Subscribe to progress updates

output <- powRICLPM(
  
  target_power = 0.8,
  
  search_lower = 100, #sample size min
  search_upper = 400, #sample size max
  search_step = 50, # step increase
  
  #time_points = c(3, 4, 5),
  time_points = 5, # timepoints used
  
  #taken from parameter definition above
  ICC = ICC,
  RI_cor = RI_cor, 
  Phi = Phi,
  
  within_cor = 0.3, #default medium size
  
  reps = 1000 #number of monte carlo simulations
  
)

})

# Revert back to sequential execution of code 
plan(sequential)

```
## 3. inspect output

```{r output}

output

summary(output)

# cross-lagged effect in different simulations
summary(output, parameter = "wB2~wA1") 
plot(output, parameter = "wB2~wA1")

```

# Power in linear mixed model (task 1.1)

```{r setup power 2}
require(lme4)
require(lmerTest)
require(simr)

```

## define parameters

```{r define parameters LMM}

#full simulation of data - linear mixed model
n_size<-90
k_size<-100 #number of trials --> auditory oddball: 
subj <- factor(1:n_size) 
trial_id <- 1:k_size #remains integer variable
condition <- c("standard", "oddball")


subj_full <- rep(subj, k_size)
trial_full <- rep(trial_id, each=n_size)
oddball_likelihood<-0.2
condition_full <- rep(condition, times=c(k_size*n_size*(1-oddball_likelihood),k_size*n_size*oddball_likelihood))
#group_full <- rep(group, k_size*n_size/2)

covars <- data.frame(id=subj_full, trial=trial_full, condition=condition_full)

#define effect parameters based on literature or experience
#-->exracted from visuospatial_JCCP and fasttask models

#FIXED EFFECTS: intercept, condition
fixed <- c(0.1, 0.3)
### --> interested fixed effect of condition d = 0.3

## Random intercepts
rand <- list(0.1, 0.1)

## residual variance
res <- 1

```

## create a model

```{r create model}

#CREATE MODEL
model <- makeLmer(y ~ condition + (1|id) + (1|trial), fixef=fixed, VarCorr=rand, sigma=res, data=covars)

summary(model)

```

## simulate model

```{r simulate LMM}

##SIMULATE
#sim_treat <- powerSim(model, nsim=10, test = fcompare(y~condition*group),lmerTestType=3)
sim_treat <- powerSim(model, nsim=1000, test = fixed("condition"),lmerTestType=3)
###--> to test a fixed effect, use test "fixed" - to test an interaction use "fcompare"

#inspect simulation
sim_treat

#debugging
# ps <- lastResult()
# ps$errors
#?powerCurve

```

# Power in linear model (task 2.1)

- simulation loop

```{r parameters linear model}

#effect size to test
a = 0.3  #desired slope - partial correlation of continuous predictor

b = 0.5  #estimated intercept - standardized
sd = 5  #estimated variability defined by standard deviation -standardized
nsim = 1000  # numerb of  simulations

pval = numeric(nsim)  #placeholder for the second for loop output
Nvec = seq(100, 200, by = 10)  #vector for the range of sample sizes to be tested
power.N = numeric(length(Nvec))   #create placeholder for first for loop output

for (j in 1:length(Nvec)) {
  N = Nvec[j]  
  x = seq(1, 10, length = Nvec[j])  #x value length needs to match sample size (Nvec) length
  
  for (i in 1:nsim) {   #for this value of N, create random error 1000 times
    y_det = a + b * x
    y = rnorm(N, mean = y_det, sd = sd)
    m = lm(y ~ x)
    pval[i] = coef(summary(m))["x", "Pr(>|t|)"]  #all the p values for 1000 sims
    
  }  #cycle through all N values
  power.N[j] = sum(pval < 0.05)/nsim  #the proportion of correct p-values (i.e the power)
}

power.N
plot(Nvec, power.N)  #need about 120 samples for 80% power

# power for n = 140 participants
power.N[4]

```

# Power in structural equation model (task 2.2)

- see file:
  - pwrSEM_model_specification.txt
  - pwrSEM_parameter_values.xlsx
-use [shinyapp pwrSEM](https://yilinandrewang.shinyapps.io/pwrSEM/)

# Power in bivaraite latent growth curve model (task 3.2)

- see file:
  - pwrSEM_GROWTH_model_specification.txt
  - pwrSEM_GROWTH_parameter_values.xlsx
-use [shinyapp pwrSEM](https://yilinandrewang.shinyapps.io/pwrSEM/)

